\documentclass[11pt,a4paper,openany]{book}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\onehalfspacing
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs,longtable,multirow,array,tabularx}
\usepackage{enumitem}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc,shapes.geometric}
\usepackage{tcolorbox}
\usepackage{titlesec}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=green!40!black,
  urlcolor=blue!70!black
}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  showstringspaces=false
}

\pagestyle{fancy}
\fancyhf{}
\lhead{Cross-Platform Research Development Assistant}
\rhead{Technical Design Monograph}
\cfoot{\thepage}
\setlength{\headheight}{14pt}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{remark}[definition]{Remark}
\newtheorem{assumption}[definition]{Assumption}

\title{\Huge\textbf{Technical Design Monograph}\\[8pt]
\Large Cross-Platform Research Development Assistant for Mathematical Finance, Macroeconomics, and Institutional Model Engineering}
\author{Prepared for Management Review}
\date{\today}

\begin{document}
\frontmatter
\maketitle
\tableofcontents
\mainmatter

%======================================================================
\chapter{Executive Summary and Recommendation}
%======================================================================

\section{Executive Summary}

This monograph proposes a shared research-development assistant for a division of approximately fifty economists, mathematical-finance researchers, and research engineers who build mathematically sophisticated, code-intensive systems for large institutional users such as central banks and financial institutions.

The proposal responds to a concrete workflow failure that is now common across modern research-engineering environments:
\begin{itemize}[nosep]
  \item paper tools such as SciSpace help with literature reading but are not code-aware;
  \item coding assistants such as Claude Code, Cursor, and Copilot help with implementation but lack durable structured research memory;
  \item manual copy-paste workflows create context loss, fragile documentation, and repeated trivial errors;
  \item there is no shared system that knows how papers, equations, code, tests, and reviewer-facing explanations fit together.
\end{itemize}

The proposed system is not an assistant replacement.  It is a \textbf{shared research intelligence layer} that sits beneath all assistants and exposes consistent research context across them.  Its core responsibilities are:
\begin{enumerate}[nosep]
  \item ingest and structure papers;
  \item maintain literature-aware research memory;
  \item connect literature, equations, code, tests, and technical documents;
  \item expose this context to Claude Code, Cursor, and Copilot;
  \item support reviewer-facing, auditable technical writing.
\end{enumerate}

The architectural recommendation is explicit:
\begin{quote}
\textbf{Build a shared backend and multiple assistant adapters, rather than a tool tied to any one AI assistant.}
\end{quote}

The first implementation should be local-first, file-first, inspectable,
and incremental.  The system should begin as a structured paper store plus
retrieval/query layer, then evolve toward richer group-wide use only after
it proves useful in real workflows.

\section{Recommendation}

The recommendation is to approve a phased build of a \textbf{Cross-Platform
Research Development Assistant} with the following strategic properties:
\begin{itemize}[nosep]
  \item assistant-agnostic core architecture;
  \item local-first and inspectable data model;
  \item explicit support for paper-to-code traceability;
  \item rigorous reviewer-oriented documentation workflows;
  \item bounded MVP scope with strong immediate value;
  \item deliberate engineering simplicity in the first version.
\end{itemize}

The rest of this monograph justifies that recommendation in technical,
operational, and organizational detail.

%======================================================================
\chapter{Problem Definition}
%======================================================================

\section{The actual problem is not ``paper search'' or ``AI coding''}

The division's workflow challenge is often misdescribed as a search or
summarization problem.  That framing is too weak.  The actual problem is
better stated as follows:
\begin{quote}
How can a research-engineering organization maintain a shared, auditable,
code-aware understanding of the literature, equations, implementation
choices, failure modes, and documentation obligations behind mathematically
sophisticated institutional software?
\end{quote}

This is a different problem from simply:
\begin{itemize}[nosep]
  \item reading papers,
  \item writing code,
  \item or using a chat assistant.
\end{itemize}

It is a systems problem in research memory, traceability, assistant
interoperability, and reviewer confidence.

\section{Why the existing tool stack is insufficient}

\subsection{Paper tools}

Tools such as SciSpace are useful for reading, annotation, and local Q\&A
about papers.  But they do not deeply integrate with:
\begin{itemize}[nosep]
  \item repository structure,
  \item implementation status,
  \item test coverage,
  \item monograph chapters,
  \item production failure modes,
  \item or equation-to-code traceability.
\end{itemize}

\subsection{Coding assistants}

Claude Code, Cursor, and Copilot are useful for implementation assistance.
But absent a structured research backend, they lack:
\begin{itemize}[nosep]
  \item durable project-level literature memory,
  \item validated citation support,
  \item paper-to-code linkage,
  \item reviewer-facing technical synthesis,
  \item and a shared state across assistants.
\end{itemize}

\subsection{Manual notes}

Private notes, copied summaries, and ad hoc documents do not scale to a
team of fifty.  They also do not provide a reliable machine-readable layer
that assistants can query.

\section{Field-specific severity of the problem}

This is especially serious in mathematical finance, macroeconomics, DSGE
modeling, and structural estimation because work in these fields combines:
\begin{itemize}[nosep]
  \item mathematical derivation,
  \item econometric assumptions,
  \item numerical methods,
  \item simulation and estimation,
  \item code implementation,
  \item and skeptical technical review.
\end{itemize}

A tool that ignores any one of those layers becomes much less useful.

%======================================================================
\chapter{Design Goals and Non-Goals}
%======================================================================

\section{Primary design goals}

The system should:
\begin{enumerate}[nosep]
  \item preserve structured research memory;
  \item connect papers, equations, code, tests, and documentation;
  \item work across Claude Code, Cursor, and Copilot;
  \item remain auditable by researchers and engineers;
  \item support skeptical reviewer-facing output generation;
  \item reduce repeated context reconstruction;
  \item improve maintainability of mathematically sophisticated production code.
\end{enumerate}

\section{Non-goals for the initial phases}

The first versions should \emph{not} attempt to be:
\begin{itemize}[nosep]
  \item a full enterprise knowledge-management platform,
  \item a perfect paper-understanding engine,
  \item a general-purpose theorem prover,
  \item a replacement for all assistants,
  \item a multi-user web product from day one,
  \item or a system that requires heavy infrastructure before delivering value.
\end{itemize}

These non-goals are important because they constrain the initial scope and
improve feasibility.

%======================================================================
\chapter{System Overview}
%======================================================================

\section{One-sentence architectural summary}

The proposed system is a \textbf{shared research intelligence backend} with
multiple thin assistant adapters.

\section{Core architectural principle}

The most important architectural decision is the separation:
\begin{equation*}
\text{shared research backend} + \text{assistant-specific adapters}
\end{equation*}
This is the only design that is robust to the division's multi-assistant
environment.

\section{High-level layered architecture}

\begin{center}
\begin{tikzpicture}[
  node distance=0.6cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=12cm, minimum height=1.0cm, align=center},
  smallbox/.style={draw, rounded corners, fill=green!5, minimum width=3.4cm, minimum height=0.9cm, align=center}
]

\node[box] (assistants) {\textbf{Assistant Integration Layer}\\ Claude Code \quad Cursor \quad Copilot \quad CLI \quad Optional UI};
\node[box, below=of assistants] (query) {\textbf{Query and Retrieval Layer}\\ paper lookup, citation graph, claim support audit, paper-to-code links, reviewer synthesis};
\node[box, below=of query] (store) {\textbf{Structured Research Store}\\ metadata, extracted text, summaries, links, reviewer notes, project overlays};
\node[box, below=of store] (ingest) {\textbf{Paper Ingestion Layer}\\ PDF extraction, metadata resolution, DOI/arXiv/OpenAlex/Crossref, text normalization};

\draw[-{Latex[length=3mm]}] (assistants) -- (query);
\draw[-{Latex[length=3mm]}] (query) -- (store);
\draw[-{Latex[length=3mm]}] (ingest) -- (store);
\end{tikzpicture}
\end{center}

\section{System context diagram}

The following diagram shows the system boundary.  The important point is
that neither Claude Code, Cursor, nor Copilot owns the canonical research
state.  They query a shared backend.  This is what makes the design usable
across tools and robust to assistant changes.

\begin{center}
\begin{tikzpicture}[
  node distance=0.9cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=3.4cm, minimum height=0.9cm, align=center},
  ext/.style={draw, rounded corners, fill=gray!12, minimum width=3.4cm, minimum height=0.9cm, align=center},
  store/.style={draw, rounded corners, fill=green!8, minimum width=4.2cm, minimum height=0.9cm, align=center}
]
\node[ext] (researcher) {Researcher /\\ Engineer};
\node[box, right=1.4cm of researcher] (backend) {Research Assistant\\ Backend};
\node[store, below=0.9cm of backend] (store2) {Structured Research\\ Store};
\node[ext, above right=0.4cm and 1.4cm of backend] (sources) {Public Literature\\ APIs and PDFs};
\node[ext, below right=0.4cm and 1.4cm of backend] (ides) {Claude Code\\ Cursor\\ Copilot};

\draw[-{Latex[length=2.5mm]}] (researcher) -- node[above]{queries / curation} (backend);
\draw[-{Latex[length=2.5mm]}] (backend) -- node[left]{read/write} (store2);
\draw[-{Latex[length=2.5mm]}] (sources) -- node[above]{metadata / text} (backend);
\draw[-{Latex[length=2.5mm]}] (ides) -- node[below]{adapter calls} (backend);
\draw[-{Latex[length=2.5mm]}] (backend) -- (ides);
\end{tikzpicture}
\end{center}

\section{Primary data flow}

A typical request flows as follows:
\begin{enumerate}[nosep]
  \item An assistant or user asks a research-aware question.
  \item The adapter converts the request into a backend query.
  \item The backend retrieves structured paper summaries, code links, and metadata.
  \item The backend returns structured context rather than free-form hidden state.
  \item The assistant uses that context to write code, documentation, or an audit.
\end{enumerate}

\section{Why this architecture is the right one}

This architecture is chosen because it directly addresses the main risks:
\begin{itemize}[nosep]
  \item \textbf{Assistant lock-in risk}: removed by separating backend from assistants.
  \item \textbf{Opaque state risk}: reduced by keeping the research store inspectable.
  \item \textbf{Cross-platform integration risk}: controlled by using multiple adapters.
  \item \textbf{Maintenance risk}: reduced by keeping the backend narrow and explicit.
\end{itemize}

%======================================================================
\chapter{Detailed Component Design}
%======================================================================

\section{Component inventory}

The system should initially consist of the following components:
\begin{enumerate}[nosep]
  \item \texttt{paper\_ingest}: import PDF, DOI, URL, arXiv.
  \item \texttt{metadata\_resolve}: query OpenAlex, Crossref, arXiv.
  \item \texttt{text\_extract}: PDF-to-text extraction and normalization.
  \item \texttt{summary\_builder}: generate structured summary records.
  \item \texttt{link\_manager}: maintain paper-to-code and paper-to-doc links.
  \item \texttt{query\_engine}: answer retrieval queries.
  \item \texttt{audit\_engine}: citation support and claim validation.
  \item \texttt{assistant\_adapters}: MCP, CLI, file-based, optional HTTP.
\end{enumerate}

\section{Proposed repository layout}

The file layout below is intended to be specific enough that an engineer or
Claude Code session could begin implementation directly from it.  The
design separates source code, research data, generated artifacts, and tests.

\begin{lstlisting}
research-assistant/
  README.md
  pyproject.toml
  requirements.txt
  .gitignore
  src/
    research_assistant/
      __init__.py
      config.py
      paths.py
      errors.py
      logging.py
      ingest/
        __init__.py
        pdf_extract.py
        metadata_resolve.py
        normalize_text.py
        deduplicate.py
        source_manifest.py
      schemas/
        __init__.py
        paper_record.py
        link_record.py
        audit_record.py
        review_record.py
      storage/
        __init__.py
        file_store.py
        index_store.py
        cache_store.py
      summarize/
        __init__.py
        draft_summary.py
        claim_support.py
        reviewer_notes.py
        synthesis.py
      links/
        __init__.py
        paper_code_links.py
        paper_doc_links.py
        equation_links.py
      query/
        __init__.py
        paper_lookup.py
        citation_graph.py
        related_work.py
        search_index.py
      adapters/
        __init__.py
        mcp_server.py
        cli.py
        workspace_exports.py
        http_api.py
      tests/
        test_ingest.py
        test_metadata.py
        test_query.py
        test_claim_support.py
        test_paths.py
        test_links.py
        test_adapters.py
  research/
    papers/
      raw/
      extracted/
    metadata/
    summaries/
    links/
    reviews/
    schemas/
    indices/
    caches/
  docs/
    design/
    usage/
    schemas/
    runbooks/
\end{lstlisting}

\subsection{File-by-file implementation intent}

\begin{longtable}{@{}p{4.0cm}p{9.5cm}@{}}
\toprule
\textbf{File or module} & \textbf{Intended responsibility} \\
\midrule
\texttt{config.py} & Global settings, environment resolution, path roots, optional API keys and local service configuration. \\
\texttt{paths.py} & Canonical path normalization across Linux, Windows, and WSL2. \\
\texttt{errors.py} & Typed exceptions for ingestion, metadata, adapter, and audit failures. \\
\texttt{pdf\_extract.py} & PDF to text extraction, fallbacks, extraction status, and provenance capture. \\
\texttt{metadata\_resolve.py} & OpenAlex / Crossref / arXiv resolution and metadata reconciliation. \\
\texttt{normalize\_text.py} & Text cleanup, whitespace normalization, header/footer stripping, and chunking support. \\
\texttt{deduplicate.py} & Duplicate DOI/title/arXiv detection and canonical paper ID assignment. \\
\texttt{source\_manifest.py} & Record the source origin, fetch date, and ingestion provenance of each paper. \\
\texttt{paper\_record.py} & Validation and serialization for paper summary schema. \\
\texttt{link\_record.py} & Validation and serialization for paper-to-code and paper-to-doc links. \\
\texttt{audit\_record.py} & Validation and serialization for claim-support audit results. \\
\texttt{review\_record.py} & Structured reviewer-facing synthesis and rebuttal notes. \\
\texttt{file\_store.py} & Read/write operations for file-based research store. \\
\texttt{index\_store.py} & Lightweight indices for search and lookup. \\
\texttt{cache\_store.py} & Cache external metadata/API responses with provenance and expiry policy. \\
\texttt{draft\_summary.py} & Generate first-pass structured summaries from extracted text and metadata. \\
\texttt{claim\_support.py} & Determine whether a citation directly, indirectly, or weakly supports a claim. \\
\texttt{reviewer\_notes.py} & Store reviewer objections, limitations, and caveats for a paper or method. \\
\texttt{synthesis.py} & Produce literature review drafts and structured synthesis across multiple papers. \\
\texttt{paper\_code\_links.py} & Create and update explicit mappings from papers to repository files and modules. \\
\texttt{paper\_doc\_links.py} & Map papers to LaTeX/Markdown documentation sections. \\
\texttt{equation\_links.py} & Optional mapping from equations or method identifiers to code or docs. \\
\texttt{paper\_lookup.py} & Query papers by title, DOI, arXiv, or semantic topic. \\
\texttt{citation\_graph.py} & Retrieve citations, cited-by relationships, and neighboring literature. \\
\texttt{related\_work.py} & Group papers by method family or implementation relevance. \\
\texttt{search\_index.py} & Search over extracted text, summaries, and notes. \\
\texttt{mcp\_server.py} & MCP tool definitions and request routing for Claude Code or compatible tools. \\
\texttt{cli.py} & Command-line entry points usable from any environment. \\
\texttt{workspace\_exports.py} & Generate assistant-agnostic context files in the repo/workspace. \\
\texttt{http\_api.py} & Optional local HTTP service for future integrations. \\
\bottomrule
\end{longtable}

\section{Data ownership by component}

\begin{center}
\begin{tabular}{@{}p{3.4cm}p{5.2cm}p{5.2cm}@{}}
\toprule
\textbf{Component} & \textbf{Writes} & \textbf{Reads} \\
\midrule
paper\_ingest & raw PDF, source manifest & URLs, DOI, arXiv IDs \\
metadata\_resolve & metadata records & source manifest \\
text\_extract & extracted text files & raw PDFs \\
summary\_builder & summary drafts & metadata + extracted text \\
link\_manager & paper/code/doc link files & summaries + repo paths \\
query\_engine & query responses (ephemeral) & all structured stores \\
audit\_engine & audit notes & summaries, citations, documents \\
assistant adapters & no canonical state by default & query layer and research store \\
\bottomrule
\end{tabular}
\end{center}

%======================================================================
\chapter{Data Model and Schemas}
%======================================================================

\section{Why explicit schemas are necessary}

A research system that relies only on free-form notes is difficult to query,
validate, and maintain.  Explicit schemas are therefore necessary to:
\begin{itemize}[nosep]
  \item reduce assistant hallucination,
  \item support consistent retrieval,
  \item allow manual inspection,
  \item and make the system auditable.
\end{itemize}

\section{Paper record schema}

A paper record should contain fields such as:

\begin{lstlisting}
{
  "id": "string",
  "title": "string",
  "authors": ["string"],
  "year": 2026,
  "doi": "string|null",
  "arxiv_id": "string|null",
  "source_url": "string",
  "abstract": "string",
  "method_family": ["string"],
  "main_contribution": "string",
  "mathematical_core": "string",
  "assumptions": ["string"],
  "exactness_status": "direct|indirect|none|unknown",
  "invertibility_mechanism": "string|null",
  "jacobian_handling": "string|null",
  "known_defects": ["string"],
  "known_remedies": ["string"],
  "scientific_relevance": "string",
  "domain_relevance": "string",
  "implementation_implications": ["string"],
  "linked_code_files": ["string"],
  "linked_doc_sections": ["string"],
  "reviewer_objections": ["string"],
  "reviewer_responses": ["string"],
  "confidence_level": "high|medium|low",
  "curation_status": "draft|reviewed|approved"
}
\end{lstlisting}

\section{Formal schema constraints}

The first implementation should also define formal validation rules.  A
compact version of the core schema constraints is:

\begin{lstlisting}
PaperRecord:
  required:
    - id
    - title
    - authors
    - year
    - source_url
    - main_contribution
    - confidence_level
    - curation_status
  constraints:
    id: nonempty string, unique
    title: nonempty string
    authors: nonempty list of strings
    year: integer in [1900, 2100]
    confidence_level: one of {high, medium, low}
    curation_status: one of {draft, reviewed, approved}

LinkRecord:
  required:
    - id
    - paper_id
    - target_type
    - target
    - relationship
  constraints:
    target_type: one of {code_file, document_section, equation, issue, test}

AuditRecord:
  required:
    - id
    - claim
    - cited_papers
    - support_classification
  constraints:
    support_classification:
      one of {direct_support, indirect_support, background_only,
              unsupported, insufficient_evidence}
\end{lstlisting}

\section{Schema evolution policy}

To keep the system stable for a team of fifty, schema evolution should be
explicitly governed:
\begin{enumerate}[nosep]
  \item schema version numbers must be stored in records;
  \item backward-incompatible changes require migration scripts;
  \item adapters must reject unknown mandatory fields unless compatibility is
        explicitly declared;
  \item changes to paper, link, or audit schemas should require review by
        both the engineer-owner and a domain reviewer.
\end{enumerate}

\section{Link record schema}

A separate link record should represent many-to-many relationships such as:
\begin{itemize}[nosep]
  \item paper $\leftrightarrow$ code file,
  \item paper $\leftrightarrow$ document section,
  \item equation $\leftrightarrow$ file/function,
  \item criticism $\leftrightarrow$ mitigation implementation.
\end{itemize}

\section{Audit record schema}

An audit record should store:
\begin{itemize}[nosep]
  \item claim text,
  \item cited paper IDs,
  \item support classification,
  \item uncertainty notes,
  \item reviewer comments,
  \item timestamp and origin.
\end{itemize}

\section{Concrete record examples}

The following examples are intentionally concrete.  They are included so
that an implementation can begin from the document without inventing record
formats later.

\subsection{Example paper record}

\begin{lstlisting}
{
  "id": "paper_hoffman_2021_neutra",
  "title": "NeuTra-lizing Bad Geometry in Hamiltonian Monte Carlo Using Neural Transport",
  "authors": ["Matthew D. Hoffman", "Pavel Sountsov", "Joshua V. Dillon"],
  "year": 2021,
  "doi": null,
  "arxiv_id": "1903.03704",
  "source_url": "https://arxiv.org/abs/1903.03704",
  "method_family": ["HMC", "transport maps", "normalizing flows"],
  "main_contribution": "Uses an invertible neural transport as a preconditioner for exact HMC in transformed coordinates.",
  "mathematical_core": "Run HMC on U(T(z)) - log|det J_T(z)| rather than on U(q) directly.",
  "exactness_status": "exact after transformed-target HMC correction",
  "invertibility_mechanism": "inverse autoregressive flow",
  "jacobian_handling": "tractable log determinant from flow architecture",
  "known_defects": ["transport objective may not align with HMC efficiency", "tail behavior must be audited"],
  "known_remedies": ["shell diagnostics", "identity initialization", "bounded scale outputs"],
  "domain_relevance": "directly relevant to learned transport for posterior geometry",
  "confidence_level": "high",
  "curation_status": "reviewed"
}
\end{lstlisting}

\subsection{Example link record}

\begin{lstlisting}
{
  "id": "link_neutra_transport_chapter",
  "paper_id": "paper_hoffman_2021_neutra",
  "target_type": "document_section",
  "target": "docs/chapters/ch16_transport_foundations.tex#sec:tf:exact_hmc",
  "relationship": "supports transformed-target HMC transport idea",
  "confidence_level": "high",
  "review_status": "reviewed"
}
\end{lstlisting}

\subsection{Example claim-support audit record}

\begin{lstlisting}
{
  "id": "audit_transport_exactness_001",
  "claim": "An imperfect but invertible transport does not bias the final HMC chain if exact HMC is run on the transformed target including the Jacobian term.",
  "cited_papers": ["paper_hoffman_2021_neutra", "paper_papamakarios_2021_flows"],
  "support_classification": "direct_support",
  "evidence_summary": "NeuTra uses learned transport followed by HMC in warped space; normalizing-flow review supports exact change-of-variables density evaluation.",
  "limitations": ["Does not guarantee improved efficiency", "Requires exact Jacobian and stable gradients"],
  "reviewer_note": "Claim is about asymptotic correctness, not performance guarantee."
}
\end{lstlisting}

%======================================================================
\chapter{Paper Ingestion and Metadata Resolution Pipeline}
%======================================================================

\section{Ingestion goals}

The ingestion layer should turn an external reference into a reproducible local research artifact.

\section{Ingestion pipeline}

\begin{center}
\begin{tikzpicture}[
  node distance=0.8cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=12cm, minimum height=0.9cm, align=center}
]
\node[box] (input) {User input: DOI / arXiv / URL / PDF};
\node[box, below=of input] (raw) {Store raw source and assign canonical local paper ID};
\node[box, below=of raw] (meta) {Resolve metadata from OpenAlex / Crossref / arXiv};
\node[box, below=of meta] (extract) {Extract PDF text and normalize};
\node[box, below=of extract] (summary) {Generate draft structured summary};
\node[box, below=of summary] (review) {Human or assistant review and curation};

\draw[-{Latex[length=3mm]}] (input) -- (raw);
\draw[-{Latex[length=3mm]}] (raw) -- (meta);
\draw[-{Latex[length=3mm]}] (meta) -- (extract);
\draw[-{Latex[length=3mm]}] (extract) -- (summary);
\draw[-{Latex[length=3mm]}] (summary) -- (review);
\end{tikzpicture}
\end{center}

\section{Ingestion pseudocode}

\begin{lstlisting}
def ingest_paper(source):
    paper_id = assign_canonical_id(source)
    raw_path = store_raw_source(source, paper_id)

    metadata = resolve_metadata(source)
    extracted_text = extract_text(raw_path)
    normalized_text = normalize_text(extracted_text)

    summary = build_draft_summary(
        paper_id=paper_id,
        metadata=metadata,
        text=normalized_text,
    )

    write_metadata(paper_id, metadata)
    write_extracted_text(paper_id, normalized_text)
    write_summary_draft(paper_id, summary)

    return paper_id
\end{lstlisting}

\subsection{Ingestion pipeline responsibilities by stage}

\begin{enumerate}[nosep]
  \item \textbf{Canonical ID assignment.}  The system assigns a stable internal paper ID before any downstream processing.  This avoids path drift when metadata later changes.
  \item \textbf{Raw source preservation.}  The original PDF or fetched representation is stored unchanged for reproducibility.
  \item \textbf{Metadata resolution.}  DOI, arXiv, title, authors, year, and source provenance are retrieved from multiple services and reconciled.
  \item \textbf{Text extraction and normalization.}  Extracted text is normalized but never treated as a replacement for the raw source.
  \item \textbf{Draft summary creation.}  A first-pass structured summary is generated and marked as draft, not authoritative.
  \item \textbf{Curation handoff.}  The paper remains in draft state until reviewed by a human or a validated assistant workflow.
\end{enumerate}

\section{Ingestion sequence diagram}

\begin{center}
\begin{tikzpicture}[
  node distance=0.55cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=11.5cm, minimum height=0.75cm, align=center}
]
\node[box] (s1) {1. Receive source: DOI / arXiv / URL / PDF};
\node[box, below=of s1] (s2) {2. Assign canonical paper ID and store raw source};
\node[box, below=of s2] (s3) {3. Resolve metadata from Crossref, arXiv, OpenAlex};
\node[box, below=of s3] (s4) {4. Extract text with status and provenance};
\node[box, below=of s4] (s5) {5. Normalize text and create searchable chunks};
\node[box, below=of s5] (s6) {6. Generate draft structured summary};
\node[box, below=of s6] (s7) {7. Run validation: schema, duplicate, extraction status};
\node[box, below=of s7] (s8) {8. Mark record as draft and await curation};
\foreach \a/\b in {s1/s2,s2/s3,s3/s4,s4/s5,s5/s6,s6/s7,s7/s8}
  \draw[-{Latex[length=2.5mm]}] (\a) -- (\b);
\end{tikzpicture}
\end{center}

\section{Failure handling in ingestion}

The ingestion pipeline must never silently pretend success.  Each stage
should return an explicit status:
\begin{itemize}[nosep]
  \item \texttt{ok},
  \item \texttt{partial},
  \item \texttt{failed},
  \item \texttt{requires\_manual\_review}.
\end{itemize}

A paper with failed PDF extraction may still have useful metadata, but it
should not be treated as fully ingested.  This distinction is important for
reviewer trust.

\subsection{Metadata reconciliation pseudocode}

\begin{lstlisting}
def resolve_metadata(source):
    candidates = []

    if has_doi(source):
        candidates.append(fetch_crossref(source.doi))
    if has_arxiv_id(source):
        candidates.append(fetch_arxiv(source.arxiv_id))

    candidates.append(fetch_openalex(source))

    merged = merge_metadata_candidates(candidates)
    validate_required_fields(merged)
    return merged
\end{lstlisting}

\subsection{Deduplication pseudocode}

\begin{lstlisting}
def assign_canonical_id(source):
    matches = []
    matches += lookup_by_doi(source)
    matches += lookup_by_arxiv(source)
    matches += lookup_by_normalized_title(source)

    if len(matches) == 1:
        return matches[0].paper_id
    if len(matches) > 1:
        return resolve_duplicate_conflict(matches)
    return mint_new_paper_id(source)
\end{lstlisting}

\section{Metadata reconciliation}

A single source should not be treated as authoritative in all cases.

Recommended preference ordering:
\begin{enumerate}[nosep]
  \item DOI/Crossref for publication metadata,
  \item arXiv for preprint metadata,
  \item OpenAlex for citation graph and broad work relationships,
  \item local human curation for final correctness if sources conflict.
\end{enumerate}

%======================================================================
\chapter{Research Store and Retrieval Layer}
%======================================================================

\section{Why file-first storage is preferred initially}

A file-first research store is preferred because it is:
\begin{itemize}[nosep]
  \item transparent,
  \item easy to version-control,
  \item robust to tool changes,
  \item debuggable by researchers,
  \item and suitable for local-first development.
\end{itemize}

\section{Retrieval operations}

The query layer should support operations such as:
\begin{itemize}[nosep]
  \item find paper by title/DOI/arXiv/topic,
  \item list citations and citing works,
  \item find papers related to a method family,
  \item read structured summary,
  \item find papers linked to code files,
  \item generate synthesis across a set of papers,
  \item audit support for a claim.
\end{itemize}

\section{Retrieval pseudocode}

\begin{lstlisting}
def find_related_papers(topic, filters=None):
    candidate_ids = search_index(topic, filters)
    papers = load_paper_records(candidate_ids)
    ranked = rank_by_relevance(papers, topic)
    return ranked


def claim_support_audit(claim, cited_paper_ids):
    papers = [load_paper_record(pid) for pid in cited_paper_ids]
    evidence = extract_relevant_support(papers, claim)
    classification = classify_support(evidence, claim)
    return {
        "claim": claim,
        "papers": cited_paper_ids,
        "evidence": evidence,
        "classification": classification,
    }
\end{lstlisting}

%======================================================================
\chapter{Assistant Integration Strategy}
%======================================================================

\section{Why multiple adapters are necessary}

Claude Code, Cursor, and Copilot differ significantly in:
\begin{itemize}[nosep]
  \item tool invocation capability,
  \item persistence model,
  \item editor integration,
  \item and how much structured context they can consume.
\end{itemize}

Therefore, the backend must expose more than one interface.

\section{Recommended access paths}

\begin{enumerate}[nosep]
  \item \textbf{MCP server} for Claude Code and any compatible environment.
  \item \textbf{CLI tools} for universal availability.
  \item \textbf{Workspace-generated files} for visibility inside all editors.
  \item \textbf{Optional HTTP API} for future richer integrations.
\end{enumerate}

\section{Illustrative MCP tool surface}

Possible MCP tools:
\begin{lstlisting}
find_paper(query)
get_paper_summary(paper_id)
papers_citing(paper_id)
papers_cited_by(paper_id)
related_papers(topic)
paper_code_links(paper_id)
claim_support_audit(claim, paper_ids)
reviewer_synthesis(topic, paper_ids)
\end{lstlisting}

\subsection{Proposed MCP tool contracts}

To reduce ambiguity, the first implementation should define a stable tool
contract.  An illustrative version is:

\begin{lstlisting}
find_paper(
  query: string,
  filters?: {
    year_min?: int,
    year_max?: int,
    method_family?: [string],
    curated_only?: bool
  }
) -> [
  {
    paper_id: string,
    title: string,
    year: int,
    relevance_score: float,
    curation_status: string
  }
]

get_paper_summary(
  paper_id: string
) -> {
  metadata: object,
  summary: object,
  links: [object]
}

claim_support_audit(
  claim: string,
  paper_ids: [string]
) -> {
  classification: string,
  evidence: [string],
  limitations: [string],
  confidence: string
}
\end{lstlisting}

\subsection{MCP implementation responsibilities}

The MCP adapter should be thin.  It should:
\begin{itemize}[nosep]
  \item validate inputs,
  \item call the backend query functions,
  \item normalize outputs into assistant-friendly structures,
  \item avoid embedding business logic that belongs in the backend.
\end{itemize}

\section{CLI equivalents}

The same operations should also exist as CLI commands so the backend remains useful outside any specific assistant.

\subsection{Illustrative CLI commands}

\begin{lstlisting}
ra paper find --query "neural transport hmc"
ra paper show --paper-id paper_hoffman_2021_neutra
ra paper citing --paper-id paper_hoffman_2021_neutra
ra audit claim --claim-file claim.txt --papers p1 p2 p3
ra links add --paper-id p1 --target src/foo.py --relationship "implements"
ra review synthesize --topic "transport maps for posterior geometry"
\end{lstlisting}

\subsection{Why CLI parity matters}

CLI parity is essential because it ensures the backend remains usable even in
assistant environments that do not support rich tool orchestration.  It also
supports scripting, CI validation, and debugging outside IDEs.

%======================================================================
\chapter{Version 2 Architecture: Multi-Parser Reading and Citation-Graph Discovery}
%======================================================================

\section{Why the architecture needs a Version 2 layer}

The personal POC revealed an important limitation: metadata APIs and simple
PDF text extraction are not enough for robust local paper ingestion.  A
Google Scholar-like experience requires two separate capabilities:
\begin{enumerate}[nosep]
  \item \textbf{strong document understanding} for papers already in the
        local library;
  \item \textbf{fast citation-graph discovery} for finding relevant papers
        beyond the local library.
\end{enumerate}

These are different subsystems and should not be conflated.  Local parsing
answers ``what does this paper say?''  Citation graph discovery answers
``which papers should I read next?''

\section{Multi-parser paper understanding subsystem}

No single open-source PDF parser is expected to be foolproof.  Therefore the
recommended design is a multi-parser consensus pipeline:

\begin{center}
\begin{tikzpicture}[
  node distance=0.55cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=11.5cm, minimum height=0.75cm, align=center}
]
\node[box] (pdf) {PDF input};
\node[box, below=of pdf] (parsers) {Run parsers in parallel: pdftotext, Marker, GROBID, MinerU, MarkItDown};
\node[box, below=of parsers] (normalize) {Normalize outputs into ParsedDocument records};
\node[box, below=of normalize] (compare) {Compare title, authors, abstract, sections, references};
\node[box, below=of compare] (reconcile) {Reconcile by agreement scoring and confidence rules};
\node[box, below=of reconcile] (store) {Store parser outputs, consensus record, and disagreement report};
\foreach \a/\b in {pdf/parsers,parsers/normalize,normalize/compare,compare/reconcile,reconcile/store}
  \draw[-{Latex[length=2.5mm]}] (\a) -- (\b);
\end{tikzpicture}
\end{center}

\subsection{Parser output schema}

Each parser should return the same abstract object:
\begin{lstlisting}
ParsedDocument:
  parser_name: string
  parser_version: string
  title_candidates: [string]
  authors: [string]
  abstract: string
  body_text: string
  body_markdown: string
  section_headings: [string]
  references: [string]
  diagnostics: object
  parse_status: ok|partial|failed
\end{lstlisting}

\subsection{Consensus output schema}

The reconciler should produce:
\begin{lstlisting}
ReconciledDocument:
  consensus_title: string|null
  consensus_authors: [string]
  consensus_abstract: string|null
  parser_agreement: object
  disagreements: [string]
  parse_confidence: high|medium|low
  requires_manual_review: bool
\end{lstlisting}

\subsection{Voting is not naive majority voting}

The system should use evidence-weighted reconciliation rather than raw
majority voting.  For example:
\begin{itemize}[nosep]
  \item title strings should be normalized and clustered by similarity;
  \item author lists should be compared by name overlap;
  \item abstracts should be checked for length and location sanity;
  \item parsers with corrupt or extremely short output should be penalized;
  \item disagreement should be preserved rather than hidden.
\end{itemize}

\section{Semantic Scholar / OpenAlex discovery subsystem}

The discovery subsystem should use external scholarly indices for breadth.
The recommended order is:
\begin{enumerate}[nosep]
  \item Semantic Scholar as the primary citation graph provider;
  \item OpenAlex as fallback and enrichment;
  \item Crossref and arXiv for metadata validation;
  \item optional later integration with Scite or Lens if institutional access
        justifies it.
\end{enumerate}

\subsection{Discovery commands}

The CLI and MCP surfaces should eventually support:
\begin{lstlisting}
ra discover --query "neural transport hmc posterior geometry"
ra citing --paper-id paper_hoffman_2021_neutra --depth 1
ra citing --paper-id paper_hoffman_2021_neutra --depth 2
ra references --paper-id paper_hoffman_2021_neutra
ra cluster-literature --seed paper_hoffman_2021_neutra --depth 2
ra survey-draft --topic "defects and remedies of NeuTra-style transport"
\end{lstlisting}

\section{Survey-generation workflow}

The intended literature-survey workflow is:
\begin{enumerate}[nosep]
  \item start from a seed paper or topic;
  \item traverse citation graph to depth 1 or 2;
  \item rank papers by relevance, citation relationship, recency, and local
        availability;
  \item ingest or parse selected papers;
  \item summarize and classify them;
  \item draft a survey organized by method family, defects, remedies, and
        applicability;
  \item keep human review as the final authority.
\end{enumerate}

\section{Why this is better than manually using Google Scholar}

Google Scholar is excellent for discovery, but it is not code-aware and is
not easily automatable.  The proposed system does not attempt to replace
Google Scholar globally.  Instead, it combines:
\begin{itemize}[nosep]
  \item Semantic Scholar / OpenAlex for citation graph traversal;
  \item local parsers for deep paper reading;
  \item structured research memory for curation;
  \item Claude Code / CLI integration for code-aware documentation.
\end{itemize}

This is the architecture most likely to help researchers read and triage
hundreds of papers while still connecting the output to implementation and
technical documentation.

%======================================================================
\chapter{Code-Aware Research Workflows}
%======================================================================

\section{Paper-to-code linking workflow}

\begin{enumerate}[nosep]
  \item researcher identifies a method paper,
  \item system retrieves or creates paper record,
  \item assistant or user proposes candidate code files,
  \item explicit links are stored,
  \item documentation can later cite both the paper and the implementation location.
\end{enumerate}

\section{Equation-to-code linking workflow}

The system should support links of the form:
\begin{itemize}[nosep]
  \item equation label → monograph section,
  \item monograph section → code file/function,
  \item code file/function → paper or method family.
\end{itemize}

This is critical for skeptical review and maintainability.

\section{Reviewer-facing synthesis workflow}

\begin{enumerate}[nosep]
  \item user asks for a literature/design section,
  \item system retrieves curated summaries,
  \item assistant generates draft synthesis,
  \item claim-support audit is run,
  \item researcher reviews and finalizes.
\end{enumerate}

%======================================================================
\chapter{Engineering Realities and Platform Risk Assessment}
%======================================================================

\section{Why this section matters}

For a research-engineering team, engineering detail is not secondary.  It is
one of the main determinants of whether a good idea becomes a reliable tool
or an expensive maintenance burden.  This section therefore surfaces the
platform and systems risks explicitly rather than minimizing them.

\section{WSL2-specific issues}

Potential issues:
\begin{itemize}[nosep]
  \item path translation between Windows host and Linux guest,
  \item slow file watching on mounted Windows filesystems,
  \item inconsistent permission behavior,
  \item intermittent DNS/network issues,
  \item tool behavior differences between files under `/home/...` and `/mnt/c/...`,
  \item IDE host/guest integration quirks.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item prefer storing active research stores under Linux-native paths,
  \item normalize all paths internally,
  \item avoid watcher-heavy architecture in the MVP,
  \item support explicit re-index commands,
  \item test WSL2 explicitly as a first-class environment.
\end{itemize}

\section{Windows-specific issues}

Potential issues:
\begin{itemize}[nosep]
  \item path separator differences,
  \item quoting and shell behavior differences,
  \item PowerShell vs bash command semantics,
  \item long path and Unicode path issues,
  \item local dependency install inconsistency.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item use Python path abstractions internally,
  \item avoid shell-specific logic in the core backend,
  \item provide CLI wrappers that are shell-agnostic where possible,
  \item document supported environment assumptions clearly.
\end{itemize}

\section{Linux-specific issues}

Potential issues:
\begin{itemize}[nosep]
  \item distro-level package differences,
  \item missing system utilities,
  \item user-level installation drift across machines.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item keep core dependencies minimal,
  \item prefer Python-based implementation where practical,
  \item document a reproducible setup process.
\end{itemize}

\section{Python environment fragmentation}

Potential issues:
\begin{itemize}[nosep]
  \item conda vs system python vs venv,
  \item package-version conflicts,
  \item institution-specific proxy/network restrictions,
  \item dependency drift between users.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item pin dependencies,
  \item provide one supported environment recipe,
  \item isolate the tool in its own environment,
  \item keep optional dependencies clearly separated.
\end{itemize}

\section{PDF and literature ingestion problems}

Potential issues:
\begin{itemize}[nosep]
  \item malformed PDFs,
  \item scanned PDFs requiring OCR,
  \item inconsistent metadata across sources,
  \item duplicate paper resolution,
  \item incomplete citation graph coverage,
  \item brittle Google Scholar behavior.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item keep raw PDF, extracted text, and curated summaries separate,
  \item use multiple metadata sources,
  \item prefer OpenAlex/Crossref/arXiv over brittle scraping,
  \item mark uncertainty explicitly.
\end{itemize}

\section{Assistant integration inconsistency}

Potential issues:
\begin{itemize}[nosep]
  \item Claude Code has richer tool integration than Copilot,
  \item Cursor may differ across versions/configurations,
  \item not every assistant will support the same retrieval path.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item backend must be independent,
  \item MCP is only one access path, not the only path,
  \item CLI and workspace files serve as universal fallback.
\end{itemize}

\section{Operational and maintenance risk}

Potential issues:
\begin{itemize}[nosep]
  \item schema drift,
  \item stale summaries,
  \item broken links to code after refactors,
  \item low team adoption if workflow is too heavy,
  \item unclear ownership.
\end{itemize}

Mitigation:
\begin{itemize}[nosep]
  \item explicit schemas,
  \item curation status fields,
  \item periodic stale-link audits,
  \item low-friction MVP,
  \item explicit operational ownership.
\end{itemize}

%======================================================================
\chapter{Testing and Validation Strategy}
%======================================================================

\section{Why testing matters unusually strongly here}

This system sits between literature, code, and documentation.  Silent errors
can therefore propagate into reviewer-facing documents and implementation
choices.  Testing is therefore not optional.

\section{Test categories}

\subsection{Unit tests}
- metadata parsing
- path normalization
- schema validation
- duplicate detection
- CLI command behavior

\subsection{Integration tests}
- DOI/arXiv/PDF ingestion end-to-end
- PDF extraction on known sample papers
- summary file creation
- paper-to-code link resolution
- claim-support audit generation

\subsection{Platform tests}
- Linux-native paths
- WSL2 Linux paths
- Windows path-handling assumptions where applicable

\subsection{Failure-case tests}
- malformed PDF
- missing metadata
- duplicate DOI
- bad Unicode path
- stale link to code file

\subsection{Assistant adapter tests}
- MCP tool schema correctness
- CLI output consistency
- workspace export file generation

\section{Representative fixture set}

The system should ship with a small fixture library of papers used for repeated testing:
\begin{itemize}[nosep]
  \item one clean arXiv paper,
  \item one DOI journal paper,
  \item one malformed PDF,
  \item one scanned/OCR-requiring PDF,
  \item one paper with many citations,
  \item one paper with ambiguous metadata.
\end{itemize}

\section{Subsystem failure-mode appendix summary}

The proposal should explicitly acknowledge that the main engineering risk is
not algorithmic elegance but operational brittleness.  The table below is a
compressed failure-mode register; a full implementation should keep the same
logic in a machine-readable issue or runbook format.

\begin{longtable}{@{}p{3.5cm}p{4.5cm}p{4.5cm}p{2.5cm}@{}}
\toprule
\textbf{Subsystem} & \textbf{Failure mode} & \textbf{Likely consequence} & \textbf{Mitigation class} \\
\midrule
Paper ingestion & malformed or encrypted PDF & missing or corrupted extracted text & preserve raw PDF, mark extraction status, allow manual fallback \\
Metadata resolver & conflicting DOI/OpenAlex/arXiv records & wrong title/authors/year, duplicate papers & reconciliation rules + human review path \\
Deduplication & two records for one paper or one ID for two papers & broken links and synthesis errors & canonical ID rules + duplicate conflict resolution \\
Structured summary & overconfident generated summary & misleading literature synthesis & draft status, curation, claim-support audit \\
Paper-code links & stale file paths after refactor & false traceability & periodic stale-link checks \\
Query layer & stale or incomplete index & missing search results & explicit reindex and cache invalidation \\
MCP adapter & tool schema drift & assistant integration failure & adapter conformance tests \\
CLI adapter & path or shell incompatibility & non-portable workflows & shell-agnostic command design \\
Workspace exports & stale generated context files & assistants reading old context & export versioning and refresh commands \\
Shared store & unauthorized edits or accidental corruption & trust loss in research memory & append-only logs / backups / review workflow \\
\bottomrule
\end{longtable}

\section{Platform test matrix}

A practical implementation should maintain a visible platform test matrix.
A minimum starting matrix is:

\begin{center}
\begin{tabular}{@{}p{3cm}p{4cm}p{6cm}@{}}
\toprule
\textbf{Platform} & \textbf{Execution mode} & \textbf{Critical checks} \\
\midrule
Linux & native & ingestion, metadata resolution, CLI, MCP, file permissions \\
WSL2 & Linux guest on Windows host & path normalization, file watching, DNS/network, mounted-drive behavior \\
Windows & optional direct support or constrained support statement & path quoting, PowerShell handling, workspace exports \\
Mixed IDE setup & VS Code / Cursor / Copilot clients & adapter invocation, generated context visibility, shared store access \\
\bottomrule
\end{tabular}
\end{center}

%======================================================================
\chapter{Security, Privacy, and Governance}
%======================================================================

\section{Security posture}

The default stance should be local-first and least-privilege.  Public paper
metadata can be fetched externally, but internal notes, code links, and
institution-specific project interpretations should remain under internal
control.

\section{Security and trust-boundary diagram}

\begin{center}
\begin{tikzpicture}[
  node distance=0.9cm,
  internal/.style={draw, rounded corners, fill=green!10, minimum width=4.4cm, minimum height=0.9cm, align=center},
  external/.style={draw, rounded corners, fill=red!8, minimum width=4.0cm, minimum height=0.9cm, align=center},
  boundary/.style={draw, dashed, rounded corners, inner sep=0.35cm}
]
\node[internal] (backend) {Internal Research Backend};
\node[internal, below=of backend] (store) {Structured Research Store};
\node[internal, right=2.0cm of backend] (assist) {Assistant Adapters};
\node[external, left=2.0cm of backend] (public) {Public Metadata and PDF Sources};
\node[external, below=of public] (userpdf) {User-Supplied PDFs};

\node[boundary, fit=(backend)(store)(assist), label=above:{Internal Trust Boundary}] {};

\draw[-{Latex[length=2.5mm]}] (public) -- node[above]{public metadata only} (backend);
\draw[-{Latex[length=2.5mm]}] (userpdf) -- node[below]{local ingestion only} (backend);
\draw[-{Latex[length=2.5mm]}] (backend) -- (store);
\draw[-{Latex[length=2.5mm]}] (assist) -- (backend);
\end{tikzpicture}
\end{center}

\section{Governance requirements}

The system should support:
\begin{itemize}[nosep]
  \item separation of public and internal content,
  \item explicit curation status,
  \item traceability of generated summaries,
  \item ownership of shared indices,
  \item and documented escalation paths for bugs and schema changes.
\end{itemize}

\section{Security design rules}

The first implementation should obey at least the following rules:
\begin{enumerate}[nosep]
  \item raw papers and extracted text remain local by default;
  \item only public metadata queries go to external APIs unless explicitly approved otherwise;
  \item internal code paths, notes, and review artifacts are never sent to public literature services;
  \item assistant adapters are not allowed to invent persistent state outside the backend-defined store;
  \item every generated summary must preserve provenance fields.
\end{enumerate}

%======================================================================
\chapter{Delivery Plan and Operational Ownership}
%======================================================================

\section{Phase 1: local MVP}

Deliverables:
\begin{itemize}[nosep]
  \item ingestion scripts,
  \item structured schema,
  \item paper store,
  \item metadata resolution,
  \item extracted text,
  \item summary drafts,
  \item CLI queries.
\end{itemize}

\section{Personal desktop POC chapter}

Before division-wide rollout, the most practical proof-of-concept is a
fully functional personal version for Claude Code on one desktop, with
optional use from a Cursor terminal on the same machine.  This smaller
scope is strategically useful because it reduces the hardest engineering
variables while still demonstrating the core value of the system.

\subsection{Why the personal POC is the right first build}

The personal POC removes several difficult classes of problems:
\begin{itemize}[nosep]
  \item no multi-user consistency or permissions coordination,
  \item no central shared service requirement,
  \item no mandatory Copilot parity in the first release,
  \item no organization-wide schema governance before the workflow is proven,
  \item and much simpler operational debugging.
\end{itemize}

At the same time, the POC should remain \emph{functionally complete}, so that
it demonstrates the real workflow rather than a toy subset.

\subsection{Functional scope of the personal POC}

The personal POC should support all of the following on a single machine:
\begin{enumerate}[nosep]
  \item ingest paper from PDF / DOI / arXiv / URL,
  \item extract text and resolve metadata,
  \item create a structured paper summary,
  \item create and query paper-to-code and paper-to-document links,
  \item expose retrieval through CLI,
  \item expose retrieval through Claude Code via MCP,
  \item allow Cursor-terminal usage through the same CLI,
  \item generate claim-support audits,
  \item generate reviewer-facing mini syntheses.
\end{enumerate}

This scope is deliberately narrow enough to be buildable, but broad enough to
prove the actual concept.

\subsection{POC-specific file layout}

A personal version can simplify the larger architecture by co-locating the
research store with the user's working environment:

\begin{lstlisting}
research-assistant-poc/
  README.md
  pyproject.toml
  src/
    research_assistant/
      ingest/
      schemas/
      storage/
      summarize/
      links/
      query/
      adapters/
      tests/
  local_research/
    papers/
      raw/
      extracted/
    metadata/
    summaries/
    links/
    reviews/
    indices/
  examples/
  docs/
\end{lstlisting}

\subsection{POC command surface}

The personal version should feel complete from day one.  A minimum useful
command surface is:
\begin{lstlisting}
ra ingest --pdf paper.pdf
ra ingest --doi 10.xxxx/xxxxx
ra ingest --arxiv 1903.03704
ra find --query "transport maps hmc"
ra show --paper-id paper_hoffman_2021_neutra
ra citing --paper-id paper_hoffman_2021_neutra
ra link add --paper-id p1 --target docs/monograph.tex --relationship supports
ra link add --paper-id p1 --target src/model.py --relationship implements
ra audit-claim --claim-file claim.txt --papers p1 p2
ra synthesize --topic "transport maps for posterior geometry"
\end{lstlisting}

\subsection{POC MCP surface for Claude Code}

The Claude Code integration should initially expose only the most useful
operations:
\begin{itemize}[nosep]
  \item \texttt{find\_paper}
  \item \texttt{get\_paper\_summary}
  \item \texttt{papers\_citing}
  \item \texttt{paper\_code\_links}
  \item \texttt{claim\_support\_audit}
  \item \texttt{reviewer\_synthesis}
\end{itemize}

This is enough to make the assistant meaningfully research-aware without
making the first implementation too broad.

\subsection{POC success criteria}

A personal POC should be judged successful if it can support the following
end-to-end demo:
\begin{enumerate}[nosep]
  \item ingest a real research paper,
  \item create a usable structured summary,
  \item link the paper to a code file and a monograph section,
  \item answer a Claude Code query about the paper and its implementation relevance,
  \item audit whether a paragraph is supported by citations,
  \item generate a reviewer-facing technical note.
\end{enumerate}

If that demo works well on a single machine, the case for group expansion is
substantially stronger.

\subsection{Why this POC remains aligned with the larger architecture}

The personal POC is not a throwaway prototype.  It should be built with the
same schemas, module boundaries, and retrieval interfaces planned for the
larger system.  The difference is only deployment scope and operational
complexity.

\subsection{Suggested staffing model}

A realistic initial staffing model is:
\begin{itemize}[nosep]
  \item \textbf{one primary engineer-owner} responsible for architecture, code integration, and release discipline;
  \item \textbf{one or two research-domain reviewers} who validate paper summaries, claim-support rules, and field-specific usefulness;
  \item \textbf{optional part-time platform advisor} for infrastructure or security review if group policy requires it.
\end{itemize}

This staffing model is designed to be achievable in a research division.  It does not assume a dedicated full-scale platform team.

\subsection{Operational ownership model}

The following ownership categories should be defined explicitly from the start:
\begin{center}
\begin{tabular}{@{}p{4.0cm}p{8.5cm}@{}}
\toprule
\textbf{Area} & \textbf{Recommended owner} \\
\midrule
Core backend code & primary engineer-owner \\
Schema changes & primary engineer-owner + research reviewer sign-off \\
Paper curation conventions & research reviewers \\
Dependency updates & primary engineer-owner \\
Assistant adapter compatibility & primary engineer-owner, with user feedback from each tool community \\
Support and bug triage & shared queue, but one named owner \\
\bottomrule
\end{tabular}
\end{center}

\subsection{MVP sprint-by-sprint execution plan}

A practical MVP can be broken into short implementation sprints.

\paragraph{Sprint 1: core paper store and ingestion.}
Implement:
\begin{itemize}[nosep]
  \item file layout,
  \item paper ID assignment,
  \item raw PDF storage,
  \item metadata resolution,
  \item text extraction,
  \item draft summary file creation.
\end{itemize}

\paragraph{Sprint 2: retrieval and CLI.}
Implement:
\begin{itemize}[nosep]
  \item paper lookup,
  \item summary retrieval,
  \item citation graph lookup,
  \item link creation,
  \item CLI commands.
\end{itemize}

\paragraph{Sprint 3: audit and review workflows.}
Implement:
\begin{itemize}[nosep]
  \item claim-support audit,
  \item reviewer synthesis generation,
  \item curation-status transitions,
  \item stale-link check.
\end{itemize}

\paragraph{Sprint 4: Claude Code adapter and workspace exports.}
Implement:
\begin{itemize}[nosep]
  \item MCP tools,
  \item workspace export files,
  \item integration tests.
\end{itemize}

\paragraph{Sprint 5: cross-tool validation.}
Validate:
\begin{itemize}[nosep]
  \item Cursor workflow,
  \item Copilot workflow via CLI/workspace files,
  \item WSL2/Linux path handling,
  \item user onboarding instructions.
\end{itemize}

\section{Phase 2: assistant integration}

Deliverables:
\begin{itemize}[nosep]
  \item MCP server,
  \item CLI adapter for all environments,
  \item workspace export files,
  \item minimal shared conventions.
\end{itemize}

\section{Phase 3: review and audit workflows}

Deliverables:
\begin{itemize}[nosep]
  \item claim-support audit engine,
  \item paper-to-code traceability workflows,
  \item reviewer synthesis generation,
  \item stale-link checks.
\end{itemize}

\section{Phase 4: group scaling}

Deliverables:
\begin{itemize}[nosep]
  \item shared repository templates,
  \item optional central index,
  \item onboarding documentation,
  \item optional web UI.
\end{itemize}

\chapter{Version 2 Roadmap: Multi-Parser Reading and Citation Discovery}
%======================================================================

\section{Why the architecture needs a Version 2 layer}

The personal POC revealed a key limitation: simple PDF text extraction plus a
single metadata API result is not reliable enough for research-grade paper
understanding.  At the same time, manual literature graph traversal is too
painful for large survey work.  Therefore the next version should explicitly
separate two problems:
\begin{enumerate}[nosep]
  \item \textbf{document understanding}: what does this paper say?
  \item \textbf{literature discovery}: which papers should I read next?
\end{enumerate}

These are different subsystems and should be engineered separately.

\section{Subsystem A: multi-parser document understanding}

No single open-source parser is reliable enough for all scholarly PDFs.  The
recommended architecture is therefore a multi-parser pipeline:

\begin{center}
\begin{tikzpicture}[
  node distance=0.55cm,
  box/.style={draw, rounded corners, fill=blue!5, minimum width=11.5cm, minimum height=0.75cm, align=center}
]
\node[box] (pdf) {PDF input};
\node[box, below=of pdf] (parsers) {Run parsers in parallel: pdftotext, Marker, GROBID, MinerU, optional MarkItDown};
\node[box, below=of parsers] (normalize) {Normalize outputs into ParsedDocument records};
\node[box, below=of normalize] (compare) {Compare title, authors, abstract, sections, references};
\node[box, below=of compare] (reconcile) {Reconcile by agreement scoring and confidence rules};
\node[box, below=of reconcile] (store) {Store parser outputs, consensus record, and disagreement report};
\foreach \a/\b in {pdf/parsers,parsers/normalize,normalize/compare,compare/reconcile,reconcile/store}
  \draw[-{Latex[length=2.5mm]}] (\a) -- (\b);
\end{tikzpicture}
\end{center}

\subsection{Design rationale}

This subsystem is needed because scholarly PDFs are not semantic documents in
any simple sense.  They are layout-heavy artifacts, and different tools are
stronger on different aspects:
\begin{itemize}[nosep]
  \item some are better at plain text extraction,
  \item some are better at scholarly structure,
  \item some are better at OCR or scanned pages,
  \item some are better at formulas, figures, or tables.
\end{itemize}

Instead of trusting one parser, the system should compare multiple parser
outputs and surface agreement, disagreement, and confidence.

\subsection{Initial parser candidates}

The planned open-source parser set is:
\begin{itemize}[nosep]
  \item \texttt{pdftotext} as simple fallback,
  \item Marker,
  \item GROBID,
  \item MinerU,
  \item optional MarkItDown as an auxiliary low-fidelity signal.
\end{itemize}

\subsection{Parser output contract}

Every parser should produce a common object, e.g.:
\begin{lstlisting}
ParsedDocument:
  parser_name: string
  parser_version: string
  title_candidates: [string]
  authors: [string]
  abstract: string
  body_text: string
  body_markdown: string
  section_headings: [string]
  references: [string]
  diagnostics: object
  parse_status: ok|partial|failed
\end{lstlisting}

\subsection{Reconciled output contract}

The reconciliation layer should produce:
\begin{lstlisting}
ReconciledDocument:
  consensus_title: string|null
  consensus_authors: [string]
  consensus_abstract: string|null
  parser_agreement: object
  disagreements: [string]
  parse_confidence: high|medium|low
  requires_manual_review: bool
\end{lstlisting}

\subsection{Important design rule}

This should not be naive majority voting on strings.  It should be
normalization + clustering + weighted agreement scoring.  The point is not to
hide disagreement; it is to make disagreement useful.

\section{Subsystem B: citation-graph discovery engine}

The second major addition is a discovery subsystem for fast literature search
and graph traversal.

\subsection{Primary and secondary discovery backends}

Recommended design:
\begin{itemize}[nosep]
  \item \textbf{Primary}: Semantic Scholar for citation graph traversal and related works.
  \item \textbf{Secondary}: OpenAlex for fallback, enrichment, and cross-checking.
  \item \textbf{Metadata support}: Crossref and arXiv where needed.
\end{itemize}

\subsection{Why this matters}

The point is to reduce painful manual workflows such as:
- finding all papers citing a seed paper,
- finding papers citing those papers,
- triaging the resulting literature,
- and extracting the subset that matters for a survey or implementation decision.

\subsection{Illustrative discovery commands}

\begin{lstlisting}
ra discover --query "neural transport hmc posterior geometry"
ra citing --paper-id paper_hoffman_2021_neutra --depth 1
ra citing --paper-id paper_hoffman_2021_neutra --depth 2
ra references --paper-id paper_hoffman_2021_neutra
ra cluster-literature --seed paper_hoffman_2021_neutra --depth 2
ra survey-draft --topic "defects and remedies of NeuTra-style transport"
\end{lstlisting}

\section{Subsystem C: download, classify, and organize workflow}

The third addition addresses a practical pain point: downloading, renaming,
and organizing papers into a local research library.

\subsection{Recommended workflow}

\begin{enumerate}[nosep]
  \item discover papers from a query or citation graph,
  \item retrieve open versions where legally available,
  \item download them into an inbox,
  \item classify and propose canonical filename/folder/tags,
  \item require user review before moving into the main library.
\end{enumerate}

\subsection{Why approval is required}

The tool should propose, not silently reorganize the library.  This is a
trustworthy research workflow, not an uncontrolled file automation system.

\subsection{Illustrative commands}

\begin{lstlisting}
ra discover-download --seed paper_hoffman_2021_neutra --depth 2 --max-papers 100
ra organize-inbox ~/GoogleDrivePapers/_inbox
ra review-proposals
ra apply-proposals
\end{lstlisting}

\section{Expected impact on literature survey workflows}

If built well, the v2 system should allow a researcher to:
\begin{enumerate}[nosep]
  \item start from a seed paper or topic,
  \item traverse the citation graph,
  \item ingest selected papers,
  \item parse and summarize them,
  \item classify them by contribution or criticism,
  \item and draft a literature survey or review-oriented memo.
\end{enumerate}

This is the direct path from today’s painful manual search process to a much
more scalable AI-assisted literature workflow.

\section{Why this remains aligned with the original architecture}

These additions do not replace the original backend.  They extend it.
The core design remains:
\begin{itemize}[nosep]
  \item assistant-agnostic backend,
  \item structured research store,
  \item explicit schemas,
  \item local-first operation,
  \item and thin adapters for assistants.
\end{itemize}

The Version 2 additions should therefore be understood as:
\begin{itemize}[nosep]
  \item parser subsystem,
  \item discovery subsystem,
  \item and paper-organization subsystem,
\end{itemize}
all attached to the same research memory foundation.


\section{Ownership model}

At minimum, the system needs explicit ownership for:
\begin{itemize}[nosep]
  \item schema changes,
  \item dependency updates,
  \item ingestion failures,
  \item integration bugs,
  \item and onboarding/support.
\end{itemize}

%======================================================================
\chapter{Anticipated Skeptical Questions and Preemptive Answers}
%======================================================================

\section{Why not just use SciSpace?}

Because SciSpace is paper-aware but not code-aware.  It does not know our
repositories, implementation files, tests, or reviewer-facing technical
traceability obligations.

\section{Why not just use Claude Code or Cursor better?}

Because the missing layer is not only better prompting.  The missing layer is durable, structured, shared research memory.

\section{Why should this be cross-platform?}

Because the division already uses multiple assistants, and a single-assistant design would reduce adoption and create lock-in.

\section{Can this team realistically build it?}

Yes, if the first version is deliberately simple:
\begin{itemize}[nosep]
  \item file-first,
  \item schema-driven,
  \item backend-plus-adapters,
  \item local-first,
  \item phased delivery,
  \item narrow MVP.
\end{itemize}
The proposal is intentionally designed to reduce platform complexity early, not increase it.

\section{Why should management believe this is not vaporware?}

Because the proposal deliberately exposes implementation details, platform risks, module boundaries, file layout, testing strategy, and phased delivery constraints.  It is designed to be close to implementation, not merely visionary.

\section{How Claude Code could implement from this document}

This document is intentionally written at a level where Claude Code or a human engineer could begin implementation directly from it.  Specifically, the document now contains:
\begin{itemize}[nosep]
  \item a repository/module tree,
  \item file-by-file responsibilities,
  \item formal record constraints,
  \item MCP and CLI contracts,
  \item ingestion pseudocode,
  \item metadata reconciliation and deduplication pseudocode,
  \item subsystem failure modes,
  \item platform test matrix,
  \item and a sprint-by-sprint MVP sequence.
\end{itemize}

The intended usage is not blind code generation from prose.  Rather, the proposal is sufficiently structured that a coding assistant can implement one subsystem at a time without inventing the architecture on the fly.

\subsection{Suggested implementation appendix workflow}

A practical execution pattern using Claude Code would be:
\begin{enumerate}[nosep]
  \item implement schemas and file-store primitives first;
  \item implement ingestion and metadata resolution second;
  \item add CLI retrieval commands third;
  \item add MCP adapters only after the backend is testable;
  \item validate each subsystem with fixture papers before enabling broader team use.
\end{enumerate}

This ordering matters.  It prevents assistant integrations from being built on top of unstable backend assumptions.

\chapter{Decision Memo for Management Review}
%======================================================================

\section{Decision requested}

Approve a phased build of a shared, assistant-agnostic research-development backend for the division.

\section{What approval enables}

Approval would authorize:
\begin{enumerate}[nosep]
  \item MVP implementation of the local-first backend;
  \item pilot integration with Claude Code and one additional assistant workflow;
  \item schema and retrieval layer validation on real research projects;
  \item preparation for broader team rollout if pilot value is demonstrated.
\end{enumerate}

\section{What approval does not yet commit the division to}

Approval of this proposal does \emph{not} commit the division to:
\begin{itemize}[nosep]
  \item building a full enterprise platform immediately,
  \item mandatory adoption by all teams,
  \item a web product or central service in phase 1,
  \item replacing existing assistants or paper tools.
\end{itemize}

\section{Recommended decision}

Approve the MVP phase with the explicit understanding that:
\begin{itemize}[nosep]
  \item the design is backend-first and cross-platform by construction,
  \item the MVP is deliberately narrow and auditable,
  \item platform and environment risks are recognized in advance,
  \item and continuation beyond the MVP should depend on demonstrated workflow value.
\end{itemize}

\backmatter
\end{document}
